\chapter{Exercise \arabic{excounter}}
\addtocounter{excounter}{1}
\label{ex:classes}
\label{ex:sex}

\herebedragons
\steps{Adding classes of individuals}{
\item Add the class \con{Person} to the \fhkb.
\item Add the classes \con{Sex}, \con{Male} and \con{Female}; make \con{Male} and \con{Female} subclasses of \con{Sex}.
\item Add the object property \con{hasSex}; restrict the class \person to \con{hasSex some Sex}.
\item Make two classes \con{Man} and \con{Woman}. Make \man\  \con{EquivalentTo: Person and hasSex some Male} and do the equivalent for \woman.
\item Run the reasoner. %Look to see where \man and \woman are in the class hierarchy; work out why.
\item Ask the DL query \con{Male and Female}; should this be possible? Make \con{Male} and \con{Female} disjoint, run the reasoner  and ask the query again. 
\item Ask the DL query \con{Sex and not (Male or Female)}; the answer suggests that there are other ways of being a \con{Sex} than being either \con{Male} or \con{Female}. Stop this happening by adding the covering axiom \con{Sex EquivalentTo: Male or Female}; run the reasoner and ask the query again.
\item Ask the DL query \con{Sex and not Female}; find an explanation for the results.
\item Ask the DL query \con{Man and Woman}; is it possible for a person to be both a man and a woman (according to our ontology)? Find out why this is possible or not.  
\item Ask the DL query \con{Person and  not (Man or Woman)}, that is, is it possible to be a person and neither a man nor a woman?
\item Ask the DL query \con{Person and not Man}; what do you find out? 
\item Make the object  property \con{hasSex} functional, run the reasoner and do all the queries again. 
\item Ask for an explanation of this inference.
% \item Add the restrictions \con{hasFather some Man} and \con{hasMother some Woman} to \person. Run the reasoner and ask which individuals have a mother. Also ask the DL query for which individuals have a grandmother. Ask for explanations.
% \item Test the \con{hasBloodRelation} property again. This time you should see it working. this is because we've now said everyone has a mother and everyone has a father, therefore everyone has a parent and thus an ancester. If everyone can have an ancestor,  then everyone is someone's descendent. 
}