\chapter{Exercise \arabic{excounter}}
\addtocounter{excounter}{1}
\herebedragons
\steps{Finding siblings}{
\item Add an object property \con{hasSibling} at the appropriate place to the object property hierarchy.
\item Decide whether it is symmetric  and/or transitive.
\item Add the property chain that will find siblings.
\item Run the reasoner.
\item Ask the DL query \con{hasSibling value \irds}; what's the problem?
\item Make \con{hasSibling} irreflexive, this should make it impossible for \rds to be his own brother. What happens? ask for an explanation (from a human being).
\item Add two sub-properties for \con{hasSibling}: \con{hasBrother} and \con{hasSister}. Decide on the transitivity, symmetry etc. for these properties and add an inverse property if you think it appropriate.
\item What sub-property chains do we need to make \con{hasBrother} work? Remember that we do not, as yet, know the sex of \rds and several other individuals.
\item The \con{isFatherOf} property has a range of \person; this will not determine the sex of a father's child. To fix this, make a property hierarchy of \con{hasChild}, \con{hasSon} and \con{hasDaughter}; add the appropriate domains and ranges.
\item Use an \con{EquivalentProperty} axiom to tie \con{hasChild} to an appropriate existing object property.
\item Add \con{hasSon} and \con{hasDaughter} assertions to the individuals.
\item Add sub-property chains to \con{hasBrother} and \con{hasSister} using these new properties; run the reasoner and test the answers with DL queries.
\item Inspect the object hierarchy.
}